---
title: 'Hyperparamter: Grid Serach from scratch'

author: 
  - Michael Hagmann^[hagmann@cl.uni-heidelberg.de]
  - Stefan Riezler
  
date: '`r format(Sys.Date(), "%B %d, %Y")`'

output:
  bookdown::pdf_book: default

urlcolor: blue

header-includes:
  - \usepackage{mathtools}
  - \usepackage{amsmath}
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache.lazy = FALSE,
                      dev = c("svglite", "pdf", "png"),
                      dpi = 300,
                      fig.path = 'figures/',
                      fig.keep = "high")


#added from: https://github.com/yihui/knitr-examples/blob/master/077-wrap-output.Rmd
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

```{r, message=FALSE}
library(tidyverse) # A collection of packages for data science. More about it on 
                   # www.tidyverse.com
library(magrittr)  # A package that provides pipe operators like %>%
library(lme4)      # A package 
library(glue)      # A package that provides interpolated string functions.
library(latex2exp) # A package that converts latex to plotmath expressions.library(lmerTest)
library(lmerTest)  # A package that improves inference for LMEM.
library(emmeans)   # A package that allows to conduct post-hoc tests.
library(optimx)    # A package that contains additional optimizer. 
library(data.table)
library(glue)
library(mgcv)
```


```{r function: ICC_nested, echo=FALSE}
ICC_nested <- function(model) {

  nested_terms <-  
    formula(model) %>%
    as.character(.) %>%
    str_extract_all(., "(?<=[|] ?)[[:word:]]+[/][[:word:]]+(?= ?[)])") %>%
    unlist(.) %>%
    na.omit(.) %>%
    as.vector(.) %>%
    str_trim(.)
  
  variances <- 
    as_tibble(VarCorr(model)) %>%
    select(effect    = grp,
           variance  = vcov) 
             

  rpl_effect <- function(x, str) {
  
    x <- str_trim(x)
    
    nested_hyperPar <- 
    strsplit(str, "/", fixed = TRUE)[[1]][1] %>%
    str_trim(.) %>%
    #add regx condition to prevent substring matching:
    paste0(., "($|:)")  
  
    x[grepl(nested_hyperPar, x)] <- str  
  
    return(x)
  }

  variances$effect <-    
    reduce(.x    = nested_terms,
           .f    = rpl_effect,
           .init = variances$effect,
           .dir  = "forward")

  variances %<>%
    group_by(effect) %>%
    summarize(variance = sum(variance, na.rm = TRUE)) %>%
    ungroup() %>%
    mutate(icc      = variance / sum(variance, na.rm = TRUE),
           icc_perc = icc * 100) %>%
    arrange(desc(icc))
  
  return(variances)
}

```



```{r read data, echo = FALSE}
grid_data <- 
  readRDS("reliability_gridsearch_results.rds")
```

TASK: Predict SOFA Score for liver based on training data without circular features.

LEARNER: The system consist of two components. The first component is a Feed Forward 
Neural Network which is trained to output a real valued score (called raw prediction) based
on 44 physiological status variables. This network is trained with a L2-loss function.
The second component is an ordinal regression model which is used to derive threshold values
to map the raw score to 0, 1, 2, 3 or 4. The output of this mapping i called the 
predicted sofa score and forms the basis of all further evaluations.

RESEARCH QUESTION: Determine a quantification of hyperparameter influence on 
summative evaluation metrics like corpus BLEU, Accuracy etc. 


PROCEDURE: We identified a set of hyperparameters we would like to study and 
defined a range of relevant values for each of them. We then choose 3 to 6 values 
that reasonable cover this range for each of the chosen hyperparameters and trained the
learner under each possible combination.

The chosen hyperparameter an values are:

```{r show hyperparamters, echo=FALSE}
summary(select(grid_data, -system, -accuracy))
```


We than evaluate the trained models on a test set (see circularity analysis for details) and 
calculated the accuracy for each of them. The results are briefly summarized by: 

```{r show accuracy, echo=FALSE}
summary(select(grid_data, accuracy))
```

To get a first impression let us plot univariate summarizations (mean and sd) for 
each hyperparameter:

```{r liver_rel_model_fig1,  echo=FALSE, warning=FALSE, fig.height = 11.7, fig.width = 8.3, fig.align = "center"}
data_plot <-
  grid_data      %>%
  select(-system) %>%
  pivot_longer(cols = !accuracy, 
               names_to = "parameter", 
               values_to = "parameter_value",
               values_transform = list(parameter_value = as.character)) %>%
  group_by(parameter, parameter_value) %>%
  summarize(mean_acc = mean(accuracy),
            sd_acc   = sd(accuracy)) 

ggplot(data = data_plot, 
       aes(y = mean_acc, x = factor(as.numeric(parameter_value)))) +
  theme_bw() + 
  xlab("") +
  ylab("mean accuracy") + 
  facet_wrap(vars(parameter), ncol = 2, scales = "free_x") +
  geom_point(size = .001) +
  geom_pointrange(aes(ymin = mean_acc - sd_acc, 
                      ymax = mean_acc + sd_acc))

```

Now we calculate a random effect model to decompose the variance:

```{r estimate var component with lme4}
mod_form <- 
  #remove seed, so that it is identical with residual
  glue("(1|{names(grid_data)[-c(1,2,9)]})") %>% 
    glue_collapse(., sep = " + ") %>%
    glue("accuracy ~ ", .)

var_decomp <-   
  lmer(as.formula(mod_form), 
       data = grid_data)

#model2 <- 
#  update(var_decomp, ~ . + (1 | hidden_number:hidden_size_max))
```

And summarize the results:

```{r, message=FALSE}
ICC_nested(var_decomp)
```

