---
title: "Model Distillation: Kidney SOFA Example"
author: 
  - Michael Hagmann ^[hagmann@cl.uni-heidelberg.de]
  - Stefan Riezler
  
date: '`r format(Sys.Date(), "%B %d, %Y")`'

output:
  pdf_document:
    latex_engine: xelatex
  bookdown::pdf_book: default
    
urlcolor: blue
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache.lazy = FALSE,
                      dev = c("svglite", "pdf", "png"),
                      dpi = 300,
                      fig.path = 'figures/',
                      fig.keep = "high")

#added from: https://github.com/yihui/knitr-examples/blob/master/077-wrap-output.Rmd
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})

```




```{r, message=FALSE}
library(tidyverse) # A collection of packages for data science. More about it on 
                   # www.tidyverse.com
library(magrittr)  # A package that provides pipe operators like %>%
library(mgcv)      # A package written by Simon Wood that implements his 
                   # (spline based) GAM specification.
library(glue)      # A package that provides interpolated string functions.
```


```{r function: powerSet, include=FALSE}
#When you want to reproduce the examples, you have to run this code chunk in advance!!
powerSet <- function(set) {
  
    apply(X      = rep(list(c(T,F)), length = length(set)) %>% expand.grid(.),
          MARGIN = 1,
          FUN    = function(...) set[...])
}
```



```{r function: Accuracy Measures for multiclass, include=FALSE}
accuracy <- 
  function(y, y_hat, filter = TRUE) mean((y == y_hat)[filter])

class_accuracy <-
  function(y, y_hat, classes = sort(unique(y))) 
    sapply(as.character(classes), function(c) accuracy(y, y_hat, filter = (y == as.numeric(c))))
```


```{r read data sets, include=FALSE}
train_data <- readRDS("train_set_with_sofa_predictions.rds")

test_data <- readRDS("test_set_with_sofa_predictions.rds")
```


Learner: Feed forward neural network with seven layers (number of neurons per 
layer: 128, 161, 203 , 256, 203 , 161, 128) and relu activation function. 
The network was trained with SGD (batch size 64) and learning rate .01 for 5 
epochs and a dropoutrate of .2 for inter hidden layer weights. All other 
optimizer settings are default values of pyTorch's SGD optimizer. The objective 
function was MSE. 

In a second step thresholds to turn the network output (score) 
into class predictions were learned . For this purpose a ordinal regression model 
was trained (R package "ordinal"). 

Train set size: `r nrow(train_data)` measurement points.
Test set size: `r nrow(test_data)` measurement points.


\newpage
## PERFORMANCE EVALUATION

### CIRCULAR MODEL - TRAIN SET ACCURACY (overall and per class) 

```{r accuracy measures on train set for circular teacher, echo=FALSE}
with(train_data, accuracy(target, sofa_circ))
with(train_data, class_accuracy(target, sofa_circ))
```

Confusion Matrix:
```{r, echo=FALSE}
xtabs(~ target + sofa_circ, data = train_data)
```


### CIRCULAR MODEL - TEST SET ACCURACY (overall and per class) 

```{r accuracy measures on test set for circular teacher, echo=FALSE}
with(test_data, accuracy(target, sofa_circ))
with(test_data, class_accuracy(target, sofa_circ))
```

Confusion Matrix:
```{r, echo=FALSE}
xtabs(~ target + sofa_circ, data = test_data)
```

\newpage
### NON-CIRCULAR MODEL - TRAIN SET ACCURACY (overall and per class) 

```{r accuracy measures on train set for non-circular teacher, echo=FALSE}
with(train_data, accuracy(target, sofa_nocirc))
with(train_data, class_accuracy(target, sofa_nocirc))
```

Confusion Matrix:
```{r, echo=FALSE}
xtabs(~ target + sofa_nocirc, data = train_data)
```


### NON-CIRCULAR MODEL - TEST SET ACCURACY (overall and per class) 

```{r accuracy measures on test set for non-circular teacher, echo=FALSE}
with(test_data, accuracy(target, sofa_nocirc))
with(test_data, class_accuracy(target, sofa_nocirc))
```

Confusion Matrix:
```{r, echo=FALSE}
xtabs(~ target + sofa_nocirc, data = test_data)
```





\newpage
### CORRELATION PATTERN
```{r load libraries to draw circular heatmap, include=FALSE}
library(circlize)
#to install package "ComplexHeatmap":
# install.packages("BiocManager")
# BiocManager::install("ComplexHeatmap")
library(ComplexHeatmap)
library(gridBase)
```


```{r calc correlations,include=FALSE}
method <- "pearson"

corr_matrix <-
  cbind(
    "label"        = sapply(select(test_data, amv:horovitz), 
                            cor,
                            method = method,
                            y = test_data$target),
    "circular"     = sapply(select(test_data, amv:horovitz), 
                            cor,
                            method = method,
                            y = test_data$sofa_circ),
    "non-circular" = sapply(select(test_data, amv:horovitz), 
                            cor,
                            method = method,
                            y = test_data$sofa_nocirc))
```



```{r kidney_distill_fig1 , fig.height = 4.5, fig.width = 5.5, fig.align = "center", warning=FALSE, message=FALSE, echo=FALSE}

plot.new()
circle_size = unit(1, "snpc") # snpc unit gives you a square region

pushViewport(viewport(x = 0, 
                      y = 0.5, 
                      width = circle_size, 
                      height = circle_size, 
                      just = c("left", "center")))


col_scale <- 
  colorRamp2(c(-1, 0, 1), c("#67001F", "#FFFFFF", "#053061"), space ="LAB")

par(omi = gridOMI(), new = TRUE)
par(mar = c(0.1, 0.1, 0.1, 0.1))
circos.par(gap.after = 35, cell.padding = c(0, 0, 0, 0))


  circos.heatmap(corr_matrix[, 3, drop = FALSE], 
                 col = col_scale, 
                 track.height = .1,
                 rownames.side = "outside",
                 rownames.font = 2)
                 #rownames.cex = 0.5 + .25 * rownames(corr_matrix) %in% vars)
  circos.track(track.index = 2, 
               panel.fun = function(x, y) {
                  cn = colnames(corr_matrix)[3]
                  n  = length(cn)
                  circos.text(rep(CELL_META$cell.xlim[2], n) + convert_x(1, "mm"), 
                              1:n - 0.5, cn, 
                              cex = 0.75, 
                              adj = c(0, 0.5), 
                              facing = "bending.inside",
                              niceFacing = TRUE,
                              font = 2)},
              bg.border = NA)
  
                
  circos.heatmap(corr_matrix[,2 , drop = FALSE], 
               col = col_scale, 
               track.height = .1)
  circos.track(track.index = 3,
               panel.fun = function(x, y) {
                  cn = colnames(corr_matrix)[2]
                  n = length(cn)
                  circos.text(rep(CELL_META$cell.xlim[2], n) + convert_x(1, "mm"), 
                              1:n - 0.5, cn, 
                              cex = 0.75, 
                              adj = c(0, 0.5), 
                              facing = "bending.inside",
                              niceFacing = TRUE,
                              font = 2)},
              bg.border = NA)
  
  
  circos.heatmap(corr_matrix[,1 , drop = FALSE], 
               col = col_scale, 
               track.height = .1)
  circos.track(track.index = 4,
               panel.fun = function(x, y) {
                  cn = colnames(corr_matrix)[1]
                  n = length(cn)
                  circos.text(rep(CELL_META$cell.xlim[2], n) + convert_x(1, "mm"), 
                              1:n - 0.5, cn, 
                              cex = 0.75, 
                              adj = c(0, 0.5), 
                              facing = "bending.inside",
                              niceFacing = TRUE,
                              font = 2)},
              bg.border = NA)
  
  circos.clear()

upViewport()

lgd_color <- 
  Legend(at = seq(-1, 1, .2), 
         col_fun = col_scale, 
         title_position = "topcenter", 
         title = "r",
         border = "black")
draw(lgd_color,  x = circle_size, just = "left")
```


\newpage
## NULLIFICATION
```{r train student gams which have access to all features, warning=FALSE, cache=TRUE, include=FALSE}
list_wo_circ <- 
  paste0(c(glue("s({names(test_data)[-c(10, 34, 42, 45:47)]})"), 
           names(test_data)[42]),
        collapse = " + ")

student_nocirc_all <- 
  gam(as.formula(paste0("sofa_nocirc ~ s(urine24, k = 150) + s(crea, k = 125) + ", list_wo_circ)), 
      data = test_data)

student_circ_all <- 
  gam(as.formula(paste0("sofa_circ ~ s(urine24, k = 150) + s(crea, k = 125) + ", list_wo_circ)), 
      data = test_data)
```



```{r train student gams which have access to non-circular features only, warning=FALSE, cache=TRUE, include=FALSE}
list_wo_circ <- 
  paste0(c(glue("s({names(test_data)[-c(10, 34, 42, 45:47)]})"), 
           names(test_data)[42]),
        collapse = " + ")

student_nocirc_noBili <- 
  gam(as.formula(paste0("sofa_nocirc ~ ", list_wo_circ)), 
      data = test_data)

student_circ_noBili <- 
  gam(as.formula(paste0("sofa_circ ~ ",list_wo_circ)), 
      data = test_data)
```


### TEACHER: circular, STUDENT: all features
```{r, echo=FALSE}
summary(student_circ_all)
```

\newpage
### TEACHER: circular, STUDENT: no bili
```{r, echo=FALSE}
summary(student_circ_noBili)
```

\newpage
### TEACHER: non-circular, STUDENT: all features
```{r,echo=FALSE}
summary(student_nocirc_all)
```

\newpage
### TEACHER: non-circular, STUDENT: no bili
```{r,echo=FALSE}
summary(student_nocirc_noBili)
```

\newpage
```{r select 5 most important featurs for non-circ model based on corr, include=FALSE, eval=FALSE}
corr_matrix[order(abs(corr_matrix[,"circular"]), decreasing=TRUE),][1:5,]
```

```{r, include = FALSE, results="hide"}
z <- plot(student_circ_all, select = 2)
shift_crea   <- abs(mean(z[[2]]$fit[between(z[[2]]$x, 0, 1)]))
shift_urine24 <- coef(student_circ_all)[1] - shift_crea
```


```{r kidney_distill_fig2, echo=FALSE, warning=FALSE, fig.height = 11.7, fig.width = 8.3, fig.align = "center"}
dev_nocirc   <- round(summary(student_circ_noBili)$dev.expl * 100)
dev_withcirc <- round(summary(student_circ_all)$dev.expl * 100)

par(mfrow = c(5,2),  mar=c(2,3,2,3), oma = c(1,1,1,1))
 #first row: urine24
   plot(x = 0:1000, y = seq(0,4, length.out=1001),
       xlim   = c(0,1000), 
       ylim   = c(0, 4), 
       main   = glue("Student without Creatinine and Urine Output (D\U00B2={dev_nocirc}%)"),
       ylab   = "", 
       xlab   = "",
       type   = "n")
  text(labels = "not included in student model",
       x      = mean(c(0,1000)),
       y      = mean(c(0,4)),
       adj    = .5)
  
  plot(student_circ_all, 
     se     = FALSE, 
     xlim   = c(0,1000), 
     ylim   = c(0,4), 
     main   = glue("Student with all Features (D\U00B2={dev_withcirc}%)"),
     ylab   = "", 
     xlab   = "",
     select = 1,
     shift  = shift_urine24)
   rug(unique(test_data$urine24))
  mtext("urine ouput", side = 4, line = 1)
  
  #second row: crea
   plot(x = seq(.15, 5.2, by = .01), y = seq(0, 4, length.out=506),
       xlim   = c(.15, 5.2), 
       ylim   = c(0, 4), 
       ylab   = "", 
       xlab   = "",
       type   = "n")
  text(labels = "not included in student model",
       x      = mean(c(.15, 5.2)),
       y      = mean(c(0,4)),
       adj    = .5)
  
  plot(student_circ_all, 
       ylab   = "",
       xlab   = "",
       xlim   = c(.15, 5.2), 
       ylim   = c(0, 4), 
       select = 2,
       shift = shift_crea )
    abline(h=0, lty = 3)
      rug(unique(test_data$crea))
  mtext("creatinine", side = 4, line = 1)
  
  #third row: calcium
  plot(student_circ_noBili, 
       se     = FALSE, 
       #rug    = TRUE, 
       ylab   = "",
       xlab   = "",
       ylim   = c(-1,4), 
       xlim   = c(.5, 1.6), 
       select = 30)
    abline(h=0, lty = 3)
    rug(unique(test_data$calcium))
  
  plot(student_circ_all, 
       se     = FALSE, 
       #rug    = TRUE, 
       ylab   = "",
       xlab   = "",
       ylim   = c(-1,4), 
       xlim   = c(.5, 1.6), 
       select = 32)
    abline(h=0, lty = 3)
    rug(unique(test_data$calcium))
  mtext("calcium", side = 4, line = 1)
  
  #forth row: bun
  plot(student_circ_noBili, 
       se     = FALSE, 
       #rug    = TRUE, 
       #ylab   = "", 
       xlab   = "",
       ylim   = c(-1,4),
       xlim   = c(10, 180), 
       select = 9)
    rug(unique(test_data$bun))
    abline(h=0, lty = 3)
  
  plot(student_circ_all, 
       se     = FALSE, 
       #rug    = TRUE, 
       ylab   = "", 
       #xlab   = "",
       ylim   = c(-1,4),
       xlim   = c(10, 180),
       select = 11)
    abline(h=0, lty = 3)
    rug(unique(test_data$bun))
  mtext("bun", side = 4, line = 1)
  
  #fifth row: chlorid
  plot(student_circ_noBili, 
       se     = FALSE, 
       #rug    = TRUE, 
       ylab   = "",
       #xlab   = "",
       ylim   = c(-1,4),
       xlim   = c(95,140),
       select = 29)
    abline(h=0, lty = 3)
    rug(unique(test_data$chlorid))
  
  plot(student_circ_all, 
       se     = FALSE, 
       #rug    = TRUE, 
       ylab   = "",
       xlab   = "",
       ylim   = c(-1,4),
       xlim   = c(95,140),
       select = 31)
    abline(h=0, lty = 3)
    rug(unique(test_data$chlorid))
  mtext("chlorid", side = 4, line = 1)
par(mfrow = c(1,1), mar = c(5,4,4,2))
```
